---
title: "skeleton of 715 project"
output: html_document
date: "2023-12-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

All text in human-readable format except comments should go into the report in the end.

#TO DO
Rubric 1: 
Statement of the research question:

1 dependent variable:

8 independent variable:

Goal:

Rubric 1.5:

Literature research and motivation:

Rubric 2: EDA and QC of data
```{r}
# TO DO: inspect if we have NA values
# TO DO: inspect num/cat, if num, what scale?
# TO DO: drop NA/impute NA
```


```{r}
# TO DO: EDA plots
```


```{r}
# TO DO: correlation and collinearity
```


```{r}
# TO DO: scaling
# TO DO: encode cat var
```

Rubric 3: 
Conduct a two- or multi-sample hypothesis test
Identify two variables that you want to test for associations, where
at least one is categorical. Choose variables relevant to your
research question
Choose the appropriate test (t test, ANOVA, Wilcox rank-sum,
Wilcoxon signed-rank, Kruskal-Wallis, binomial test, proportion test,
chi-square test, Fisher exact test)
Check any assumptions required and justify your choice of test
```{r}

```

Rubric 4-5:
Choose a regression or classification model and run the model
justify decision
variable selection
assumption check
report key statistics
```{r}

```

Rubric 6: 
Evaluate model fit
Choose, calculate, and appropriately interpret a metric
Justify the metric you chose based on the goal of your model
Include 1+ plot that you use to evaluate model fit
```{r}

```

Rubric 7:
Compare your model to one with different variables
can change the variables in any way that you think would be
interesting: e.g., add/remove variables, different regularization or
scaling. Justify the alternative model that you choose.
```{r}

```

Rubric 8:
Conduct one follow-up analysis
```{r}

```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
